\listofpapersintro{This thesis is based on the following papers, which are referred in the text by their roman numerals.}% 
%\listofpapersoutro{Reprints were made with permission from the publishers.}

\newcommand{\conference}[1]{\marginpar{\tikz[baseline=(c.base)]\node(c)[conference]{#1};}}

% Environment used to create a list of papers
\begin{listofpapers}
  %
  \item \label{paper:VMCAI13}%
    {\bf All for the Price of Few} %{\small (Parameterized Verification through View Abstraction)}}
    \conference{VMCAI'13}
    \\{\footnotesize Parosh A.\ Abdulla, Fr{\'{e}}d{\'{e}}ric Haziza, and Luk{\'{a}}s Hol{\'{\i}}k}.
    \\In {\it Verification, Model Checking, and Abstract Interpretation}, 2013.
    % 
  \item \label{paper:TACAS13}%
    {\bf An Integrated Specification and Verification Technique}\conference{TACAS'13}
    \\{\bf for Highly Concurrent Data Structures}
    \\{\footnotesize Parosh A.\ Abdulla, Fr{\'{e}}d{\'{e}}ric Haziza, Luk{\'{a}}s Hol{\'{\i}}k, Bengt Jonsson, and Ahmed Rezine}.
    \\In {\it Tools and Algorithms for the Construction and Analysis of Systems}, 2013.
    \\{\it\color{magenta} Best paper award} at the {Joint European Conferences on Theory and Practice of Software, {ETAPS} 2013.}
    % 
  \item \label{paper:SAS14}%
    {\bf Block Me If You Can! {\small (Context-Sensitive Parameterized Verification)}}
    \conference{SAS'14}
    \\{\footnotesize Parosh A.\ Abdulla, Fr{\'{e}}d{\'{e}}ric Haziza, and Luk{\'{a}}s Hol{\'{\i}}k}.
    \\In {\it Static Analysis Symposium}, 2014.
    % 
  \item \label{paper:CAV08}%
    {\bf Monotonic Abstraction for Programs with Dynamic Memory Heaps}
    \conference{CAV'08}
    \\{\footnotesize Parosh A.\ Abdulla, Ahmed Bouajjani, Jonathan Cederberg, Fr{\'{e}}d{\'{e}}ric Haziza, Ahmed Rezine}.
    \\In {\it Computer Aided Verification}, 2008.
    % 
  \item \label{paper:FORTE08}%
    {\bf Parameterized Tree Systems}
    \conference{FORTE'08}
    \\{\footnotesize Parosh A.\ Abdulla, Noomene Ben Henda, Giorgio Delzanno, Fr{\'{e}}d{\'{e}}ric Haziza, Ahmed Rezine}.
    \\In {\it Formal Techniques for Networked and Distributed Systems}, 2008.
    % 
  \item \label{paper:MCC08}%
    {\bf Model Checking Race-Freeness}
    \conference{SIGARCH'08}
    \\{\footnotesize Parosh A.\ Abdulla, Fr{\'{e}}d{\'{e}}ric Haziza, and Mats Kindahl}.
    \\In {\it SIGARCH Computer Architecture News}, 2008.
    % 
  \item \label{paper:STTT15}%
    {\bf Parameterized Systems through View Abstraction}
    \conference{STTT'15}
    \\{\footnotesize Parosh A.\ Abdulla, Fr{\'{e}}d{\'{e}}ric Haziza, and Luk{\'{a}}s Hol{\'{\i}}k}.
    \\In {\it International Journal on Software Tools for Technology Transfer}, 2015.
    % 
\end{listofpapers}
