\section{Implementation Issues and Experimental Data}
\label{implementation:section}

Based on the techniques described above, we have implemented a prototype in Java. We have chosen to verify null-dereferencing, wellformedness of output and garbage creation for a collection of programs taken from the PALE website. The encouraging results, obtained on a 1.1 GHz Pentium M, are summarised in the table below.

To check null dereferencing means to guarantee the abscence of executing statements of the form $\xvar\assigned\nxt\yvar$ or $\nxt\xvar\assigned\yvar$ if $\labeling(\xvar)=\labeling(\nil)$. This is checked by inspecting the program and look for all transitions $\transition=\tuple{\cstate_1,\action,\cstate_2}$ where there is an action $\action$ of the mentioned form. For all such $\transition$, we create the configuration $\conf_1=\tuple{\cstate_1,\graph}$ where $\graph$ = $\tuple{\vertices,\suc,\labeling}$ such that $\vertices=\{\vertex\}$, $\suc$ is undefined everywhere and $\labeling$ is undefined everywhere except for $\xvar$ and $\nil$, whose images are both $\vertex$. The set of these configurations are our $\finalconfs$.

Wellformedness of output means that the result of the program is a well formed list, i.e. it is null terminated, so it does not contain a cycle, and there is no node with more than one predecessor. To check this property, we create the constraints shown in the image.

Garbage means intuitively that when the program finishes, all cells are reachable from a given set of pointers. This means that the set $\finalconfs$ contains the four configurations shown in the picture.

The experimental results are encouraging, and they can be seen in the table below:

\begin{tabular}{l|c|c|c|c}
Program & Property & Time & Number of & Iterations \\
 & checked & & constraints & \\
\hline
\verb+Concatenate+ & Deref & 0.4& 7 & 3\\
\verb+Delete+ & Deref & 0.4 & 8 & 4\\
\verb+Fumble+ & Deref & 0.3 & 3 & 2\\
\verb+Reverse+ & Deref & 0.3 & 2 & 1\\
\verb+Walk+ & Deref & 0.4 & 9 & 3\\
\verb+Zip+ & Deref & 1.9 & 206 & 12\\
\verb+Fumble+ & Garbage & 0.7 & 38 & 14\\
\verb+Reverse+ & Garbage & 0.8 & 55 & 24\\
\verb+Reverse+ & Wellformedness & 1.7 & 48 & 20
\end{tabular}
