%% ====================================================================
\chapterwithtoc{Parameterized Systems}
\label{chapter:parameterized:systems}
\index{Parameterized Systems}%
%% ====================================================================
In this thesis, we focus on systems consisting of an arbitrary number
of processes. The size of the system is the parameter of the
verification problem.
% 
For each value $n$ of the parameter, the system $S_n$ is the parallel
composition of $n$ processes.
%
\index{Safety}%
\index{Initial configurations}%
\index{Bad configurations}%
\index{Reachability}%
Given a specification, we are interested in proving safety for the
family $(S_n)_{n\ge 0}$ --- as~a~whole --- which is obtained through
the non-reachability of some set \Bad\ of \emph{bad configurations},
starting from some set~\Inits\ of \emph{initial configurations}. %
%
Processes are usually, but not necessarily, copies of each other.
%
\[ S_n = \underbrace{P\ \|\ P\ \|\ \cdots\ \|\ P}_{n\text{ times}}\hspace{2cm}\text{Can } (S_n)_{n\ge 0} \text{ reach }\Bad\text{ from }\Inits\text{ ?} \]
% 

%% ********************************************************************
\KW{Examples}%
\index{Mutual Exclusion}%
\index{Distributed Protocols}%
\index{Cache Coherence Protocols}%
%% ********************************************************************
Parameterized systems arise naturally in the modeling of mutual
exclusion algorithms, distributed protocols, or cache coherence
protocols. %
For instance, sensor networks typically consist of thousands of
identical nodes, web services must handle millions of requests of the
same type. %
Mutual exclusion must be guaranteed regardless of the number of
processes that participate in a given session of the protocol. %
Cache coherence must be ensured regardless of the number of cache
lines or the number of physical processors.
% 
Those systems can be handled easily for small values of the parameter,
that is, if the system involve a few components. It is interesting to
extend the verification to the parameterized case, even though the
%total 
state-space for the whole family is infinite.
\index{Infinite-state systems}%

\input{chapters/parameterized-systems/features}
\input{chapters/parameterized-systems/formal-definition}
\input{chapters/parameterized-systems/non-atomic}
\input{chapters/parameterized-systems/reachability}
\newpage
\input{chapters/parameterized-systems/szymanski}

%% ====================================================================
\whatwelearned{parameterized-systems}
%% ====================================================================
