%% ====================================================================
\chapterwithtoc{Monotonic Abstraction}
\label{chapter:monotonic:abstraction}
\index{Abstraction}%
%% ====================================================================

We now present a technique to solve the reachability problem presented
in the previous chapter. This technique has been introduced by Abdulla
et al.~\cite{ACJT00} and is based on the Well-Quasi Ordering framework.

We first introduce the notion of ordering and monotonicity and how we
use them to derive an over-approximation. We then present the
algorithm to prove the safety property. Finally, we discuss
termination and show how we applied the method on a few protocols.
%
While the method has been applied to linear topologies~\cite{IJFCS09},
one of the contributions of this thesis is to apply it to multiset
topologies, to shape analysis and to extend it to tree topologies.

\input{chapters/monotonic-abstraction/upward-closed-sets}
\input{chapters/monotonic-abstraction/monotonicity}
\input{chapters/monotonic-abstraction/backward-reachability}
\input{chapters/monotonic-abstraction/procedure-termination}

%% ====================================================================
\section{Applications}
\index{Experiments}%
%% ====================================================================

\index{Cache Coherence Protocols}%
\index{Distributed Protocols}%
The method has been applied on a variety of protocols such as cache
coherence protocol or distributed protocols, presented in
Chapter~\ref{chapter:experimentation}.
%
In~\cite{ParamVerif:global:conditions,Parosh:non-atomic}, Abdulla et
al.\ apply the method to parameterized systems, in particular with
linear topologies and atomic global conditions. We have applied the
method to protocols modeled as a Petri Net (i.e.\ a multiset) and
extended it to tree topologies.

\input{chapters/monotonic-abstraction/applications-pn.tex}
\input{chapters/monotonic-abstraction/applications-shapes.tex}
\input{chapters/monotonic-abstraction/applications-trees.tex}


%% ====================================================================
\whatwelearned{monotonic-abstraction}
%% ====================================================================
